Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-11 Master Thesis/synopsis.tex
Update

\begin{proof}
Since any rational function on $X$ is uniquely determined by its values at those points where it is defined, $\varphi$ is unique. In other words, the following maps are injective for all finite subsets $S\subset X_0$:
\[ r_S:K^S \to\prod_{P\in X_0\setminus S} k(P)^\times,\ x\mapsto (r_P(x))_P \]
For existence, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the composition $\varrho:K'^{S'}\to\prod k(\sigma P)^\times\to \prod k(P)^\times\to\operatorname{coker}(r_{S,T})$ is trivial. Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.

@@

Note that for every $x\in K^\times$ we have $v_P(x)= 0$ for almost all $P\in X_0$; similarly for $K'$ and $X'_0$. By Lemma \ref{4.5} (i) we can find an admissible set $S\subset X_0\setminus \{P\}$ such that $x\in K'^{S'}$ for any $x\in U_{P'}$.
\end{proof}

@@

\begin{remark}
For any finite subextension $L/K$ of $\overline{K}/K$ there exists an isomorphism $\varphi_L: L'\to L$. If $K\subset L\subset M$ are finite subextensions, we have $\varphi_M|_{L'} = \varphi_{L'}$, since $r_Q(\varphi_M(x)) = r_Q(\varphi_L(x))$ for all $x\in L'^\times$ and almost all $Q\in (X_M)_0$. By passing to limits we obtain an isomorphism $\varphi: \overline{K}'\to \overline{K}$.
\end{remark}

>>>>

\begin{proof}
By Lemma \ref{5.4}, the following maps are injective for all finite subsets $S\subset X_0$:
\[ r_S:K^S \to\prod_{P\in X_0\setminus S} k(P)^\times,\ x\mapsto (r_P(x))_P \]
Since any rational function on $X$ is uniquely determined by its values at those points where it is defined, $\varphi$ is unique.

For existence, let $(S,T)$ be an admissible pair. First, we show that there is a unique homomorphism $\varphi^{S,T}$, making the following diagram commutative:
\[ (\star) \qquad \begin{tikzcd}
K'^{S'} \rar["\varphi^{S,T}", dashed] \dar["r_{S',T'}"'] & K^S \dar["r_{S,T}"]\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times \rar["\prod\varphi_P"'] & \displaystyle \prod_{P\in T}k(P)^\times
\end{tikzcd} \phantom{\qquad (\star)} \]
The vertical maps are injective by definition. Thus, it suffices to show that the composition $\varrho:K'^{S'}\to\prod k(\sigma P)^\times\to \prod k(P)^\times\to\operatorname{coker}(r_{S,T})$ is trivial. Consider the commutative diagram which results from tensoring $(\star)$ with $\mathbb{Z}/n\mathbb{Z}$:
\[ \begin{tikzcd}
K'^{S'}/n \rar[dashed]\dar & K^S/n\dar\\
\displaystyle \prod_{P\in T}k(\sigma P)^\times/n \rar & \displaystyle \prod_{P\in T}k(P)^\times/n
\end{tikzcd} \]
According to Lemma~\ref{4.3} and Lemma~\ref{4.5} (ii) this diagram can now be completed by $\varphi_n: K'^{S'}/n\to K^S/n$. This implies $\varrho/n=0$, i.e. $\operatorname{im}(\varrho)\subset n\operatorname{coker}(r_{S,T})$ for every $n\in\mathbb{N}$. By Lemma~\ref{5.2} (iii) and Lemma~\ref{5.4} it follows $\varrho = 0$. By symmetry, $\varphi^{S,T}$ is an isomorphism.

@@

The commutativity of the diagram follows from the commutativity of $(\star)$ with $T=X_0\setminus S$. Note that with Lemma \ref{4.5} (i), we can find for all $x\in U_{P'}$ an admissible set $S\subset X_0\setminus \{P\}$ such that $x\in K'^{S'}$.
\end{proof}

@@

\begin{remark}
For any finite subextension $L/K$ of $\overline{K}/K$ there exists an isomorphism $\varphi_L: L'\to L$. If $K\subset L\subset M$ are finite subextensions, we have $\varphi_M|_{L'} = \varphi_{L'}$, since $r_Q(\varphi_M(x)) = r_Q(\varphi_L(x))$ for all $x\in L'^\times$ and almost all $Q\in (X_M)_0$ according to Remark \ref{3.3} and Step 1. By passing to limits we obtain an isomorphism $\varphi: \overline{K}'\to \overline{K}$.
\end{remark}
